---
title: "Week 6: RMarkdown"
author: "Keanu Rochette-Yu Tsuen"
date: "2024-10-01"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
# setting for the whole Markdown doc when knitted
#knitr::opts_chunk$set(echo = TRUE, # show outputs all the time
#                      warning = FALSE) # Don't show warnings all the time 


knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      warning = FALSE)
```

# Introduction to RMarkdown

Let's give it a shot a *RMarkdown*!\
We will use the Penguin data to practice.

## Load the Libraries
```{r}
library(palmerpenguins)
library(tidyverse)
library(here)
library(kableExtra) # to make tables in RMarkdown
```

## Comments in RMarkdown 
Comment structure use <&!-- XXX -->, without the "&".  
<!-- This becomes a comment and will not be displayed after knitting -->
This is comment will be shown tho.

## Inspect the data
```{r}
glimpse(penguins)
```
# Figures in RMarkdown

## Make a simple plot
```{r}
ggplot(penguins,
        aes(x= bill_depth_mm, y = bill_length_mm, group = species)) +
  geom_point(aes(color = species))+ 
  geom_smooth(method = "lm")+
  labs(x= "Bill Depth (mm)",
       y = "Bill Lenghth (mm)")
```

## Changing the size
Specify those within the {r} of the code chunk.  
Those can also be specified in the knitr::opts_chunk section so it applies to all of the figures produced in the document.   
- out.width = "70%" (70% of the width)   
- fig.width = 3 (width 3 in)   
- fig.height = 4 (height 4 in)   
- fig.asp = 1.2 (aspect ratio of 1.2)  
- fig align = "center"   

#### out.width = "70%"
```{r, out.width="70%"}
ggplot(penguins,
        aes(x= bill_depth_mm, y = bill_length_mm, group = species)) +
  geom_point(aes(color = species))+ 
  geom_smooth(method = "lm")+
  labs(x= "Bill Depth (mm)",
       y = "Bill Lenghth (mm)")
```

#### fig.width = 4, fig.height= 3
```{r, fig.width = 4, fig.height= 3}

ggplot(penguins,
        aes(x= bill_depth_mm, y = bill_length_mm, group = species)) +
  geom_point(aes(color = species))+ 
  geom_smooth(method = "lm")+
  labs(x= "Bill Depth (mm)",
       y = "Bill Lenghth (mm)")
```

#### fig.asp= 1.2
```{r, fig.asp= 1.2}

ggplot(penguins,
        aes(x= bill_depth_mm, y = bill_length_mm, group = species)) +
  geom_point(aes(color = species))+ 
  geom_smooth(method = "lm")+
  labs(x= "Bill Depth (mm)",
       y = "Bill Lenghth (mm)")
```

#### fig.asp= 1.2, fig.align= "center"
```{r, fig.asp= 1.2, fig.align= "center"}

ggplot(penguins,
        aes(x= bill_depth_mm, y = bill_length_mm, group = species)) +
  geom_point(aes(color = species))+ 
  geom_smooth(method = "lm")+
  labs(x= "Bill Depth (mm)",
       y = "Bill Lenghth (mm)")
```

## Save the figures
This is an alternative to ggsave, an automated way.  
It's important that all the code chunks are *NAMED*. {r _label_}, no spaces.
You can only have *ONE figure* per R chunk.  

Set the figure path within knitr::opts_chunk.  
We can't use here() in this case unfortunately.    
--> fig.path = "../outputs/"

#### fig.path = "../outputs/"
```{r penguinplot, fig.path ="../outputs/", fig.asp= 1.2, fig.align= "center"}

ggplot(penguins,
        aes(x= bill_depth_mm, y = bill_length_mm, group = species)) +
  geom_point(aes(color = species))+ 
  geom_smooth(method = "lm")+
  labs(x= "Bill Depth (mm)",
       y = "Bill Lenghth (mm)")

# this figure was indeed saved in the outputs folder
```

## Giving the figure a caption: fig.cap = "..."
```{r, fig.asp= 1.2, fig.align= "center", fig.cap= "This is a scatter plot of penguin data"}

ggplot(penguins,
        aes(x= bill_depth_mm, y = bill_length_mm, group = species)) +
  geom_point(aes(color = species))+ 
  geom_smooth(method = "lm")+
  labs(x= "Bill Depth (mm)",
       y = "Bill Lenghth (mm)")
```

## Saving the figure in a specific format: dev = " _specific format type_ "
```{r penguinpdf, fig.asp= 1.2, fig.align= "center", fig.cap= "This is a scatter plot of penguin data", fig.path ="../outputs/", dev="pdf"}

ggplot(penguins,
        aes(x= bill_depth_mm, y = bill_length_mm, group = species)) +
  geom_point(aes(color = species))+ 
  geom_smooth(method = "lm")+
  labs(x= "Bill Depth (mm)",
       y = "Bill Lenghth (mm)")

#Note: this figure was saved as a pdf. In the html output, it shows as a scrolling figure
```

# Tables in RMarkdown
## Manual method
You can do it by hand. It is extremely tedious and takes accuracy to code it right.    

| Time  | Session | Topic    |  
|:------|:-------:|---------:|  
| _left_ | _center_| _right_ |  
| 01:00  | 1       | Anatomy |  
| 01:50  |         | *Break* |  
| 02:00  | 2       | Tables  |  
| 02:45  |         | *Break* |  

<!-- ":" shows how to the justify the column-->


## Tables with kable
Load the kableExtra package in the library section. 

```{r}
penguins %>%  
  group_by(species) %>% 
  summarize(bill_mean = mean(bill_length_mm, na.rm = T)) %>% 
  kbl() # generates a kable table
```


You can also change the look of the table by adding more details and instructions
```{r}
penguins %>%  
  group_by(species) %>% 
  summarize(bill_mean = mean(bill_length_mm, na.rm = T)) %>% 
  kbl() %>% 
  kable_classic() %>% # table theme
  row_spec(2, bold = T, color = "white", background = "red")
# select row 2, bold text, white color and red background
```


Changing the size of the table  
```{r}
penguins %>%  
  group_by(species) %>% 
  summarize(bill_mean = mean(bill_length_mm, na.rm = T)) %>% 
  kbl() %>% 
  kable_classic() %>% # table theme
  row_spec(2, bold = T, color = "white", background = "red") %>% 
      # select row 2, bold text, white color and red background
  kable_styling(full_width = F) # gives a narrower table
```

There are lot more packages that can make tables even better in RMarkdown. 
https://rfortherestofus.com/2019/11/how-to-make-beautiful-tables-in-r/


# Themes in RMarkdown
So many to choose from and more to download.

https://www.datadreaming.org/posts/2018-04-11-r-markdown-theme-gallery/2018-04-11-r-markdown-theme-gallery

# Output formats 
Try Github Document: https://rmarkdown.rstudio.com/github_document_format.html

Each type of outputs has different arguments and are not transferable. 
Beware of the type chosen.





