---
author: 'Keanu Rochette-Yu Tsuen'
date: '`r format(Sys.Date())`'
title: Week 6 - RMarkdown Homework
output: 
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
---
<!-- Changing the header color of the Cayman theme
Specifically, creating a color gradient for the banner -->
```{css my-header-colors, echo = FALSE}
.page-header {
    background-image: linear-gradient(160deg, #2980B9, #6DD5FA, #FFFFFF);
}
```


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Introduction - Taking a shot at using RMarkdown
_Disclaimer: this assignment will be (a little) chaotic. Enjoy!_  


Howdy y'all, today we'll create my first markdown document (_+ output_ ).    
Not gonna lie this one was a tough one, the table in particular was giving me a hard time.  
But in the words of Sia:  

> "I'm unstoppable. I'm a Porsche with no breaks."   
> _Sia, somewhere cool_

## Load the Libraries
To start and **_slay_** that code, we are loading them libraries. 
```{r}
library(tidyverse) #to keep things organized
library(here) # so our pathways are unbreakable
library(gt) # tables baby! 
library(ghibli) # for the pretty colors
```

## Load the Data
We need something to work with.  
We'll use data collected by the Silbiger Lab ( _yas!_ ).
```{r}
chem_data<- read_csv(here("Week_04", "data", "chemicaldata_maunalua.csv"))
# gotta use read_csv instead of read.csv, more demure
```

## Making a Summary Table
To show off those coding skills, we'll make a "simple" summary table. ( _he thought..._ )
```{r}
chem_data %>% 
    # remove those NA gurl! bad vibes
  drop_na() %>% 
    # Swith her up so she's longer
  pivot_longer (cols = Temp_in:percent_sgd,
                names_to = "Parameter",
                values_to = "Values") %>% 
    # I don't need those either
  select(-Site, -Season, -Tide_time, -Waypoint) %>% 
    # cluster the data by zone and by parameter
  group_by(Zone, Parameter) %>% 
    # then we can get our little summary
  summarize(Mean = mean(Values), Median = median(Values),
            StandDev= sd(Values)) %>% 
    # we'll round those values #cutesy
  mutate_if(is.numeric, round,digits = 2) %>% 
    # Problems start here: we need a wide boi for this task
    # We also need the columns with names as ZONE_VALUE, 
    # hence why I used the names_glue function
  pivot_wider(names_from = "Zone", 
              values_from = c(Mean:StandDev),
              names_glue = "{Zone}_{.value}") %>% 
    # I needed to rearrange my columns so all the summaries are by "Zones". 
    # But ya boi doesn't know any other way to do it. 
    # So I just had to type it by hand. Send help please...
  select(Parameter, 
         Ambient_Mean, Ambient_Median, Ambient_StandDev,
         Diffuse_Mean, Diffuse_Median, Diffuse_StandDev,
         Offshore_Mean, Offshore_Median, Offshore_StandDev,
         Transition_Mean, Transition_Median, Transition_StandDev) %>% 
    # Anyway, then we can set the biogeochem parameters as the row names.
  gt(rowname_col = "Parameter") %>% 
    # Gotta keep it nice and tidy with a title and subtitle
  tab_header(title = "Summary Statistics of Biogeochemical Parameters", 
             subtitle= "Taken in Maunalua Bay")  %>% 
    # All that hard work for the grand finalé. 
    # We are creating over arching column categories, 
    # representing the sampled zones, under which the summary statistics 
    # will be nicely grouped.
  tab_spanner_delim(delim = "_") 
```
Look at that beautiful table !  

> "Very Mindful! Very Demure!  
> _Joolie Lebron, Tiktok (I think) _

## Making a Pretty Graph
K! Now that the table is done, we can try to make a graph with ggplot.  
There are many options here.  

- time series _(actually no, we don't have any dates. LOL! )_
- box plots
- linear regressions

But today, I decide to be wild and make violin plots. 

```{r violinplot, fig.height= 12, fig.width= 10, fig.path ="../outputs/"}
chem_data %>% 
    # again, run away from problems, run away from NAs.
  drop_na() %>% 
    # Idk, the names weren't my vibe so I changed them
  rename(Nitrate_Nitrite = NN,  Percent_SGD= percent_sgd,
         Total_Alkalinity = TA, Temperature_C= Temp_in) %>%
    # Long boiiiii
  pivot_longer (cols = Temperature_C:Percent_SGD,
                names_to = "Parameter",
                values_to = "Values") %>% 
    # I didn't need those so ... gurl bye!
  select(-Site, -Season, -Tide_time, -Waypoint) %>% 
    # now the fun part, ggplot ! 
  ggplot(aes(x= Zone, y = Values, fill = Zone)) +
    # make them violin plots
  geom_violin() +
    # we be wild but tidy, everything in 2 columns only.
    # and also each graph gets their own scale, bcs idk... equity? 
  facet_wrap(~Parameter, ncol = 2, scales= "free") +
    # graph needs to be clear, labels, labels, labels
  labs(title = "Biogeochemical Parameters in Different Zones of a Lagoon",
       subtitle = "Data Collected from Maunalua Bay",
       x = "Zone sampled",
       y = "Measured Values", #give pretty names to the plot and axes
       fill = "Zones",
       caption = "Note: Data provided by the Silbiger Lab, for educational purposes only, yasss~ !") + 
    # we need something basic to balance it out: theme_bw()
  theme_bw() + 
    # yas-ifying that graph with some font size change and a background color
  theme(plot.title = element_text(size=16), 
        plot.subtitle = element_text(size=14),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12),
        plot.caption = element_text(size = 12),
        panel.background = element_rect(fill = "#FFFAF1"), 
        legend.position = "none") + # no need legend, redundant here
    # yas-ifying her even more with some ghibli theme colors
  scale_fill_ghibli_d("SpiritedMedium", direction = -1)
        
 
  
```

## Closing remarks
Thanks for reading til here!  
Hope that was entertaining and not too cringey.  
_Maybe a bit cringey tho hihi!_


