---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: Week 10 - Tidy Tuesday 1
output: 
  github_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```


# Load the libraries
```{r}
library(tidyverse)
library(multcompView)
library(ghibli)

```

# Load the data
This week we have a data set about monster movies. 
```{r}
tuesdata <- tidytuesdayR::tt_load('2024-10-29')

monster_movie_genres <- tuesdata$monster_movie_genres
monster_movies <- tuesdata$monster_movies
```

## Cleaning the Data 
I'm grouping the data by decade to create a new column variable. 
```{r}
monster_movies <- monster_movies %>% 
  mutate(decade = floor(year / 10) * 10) %>% 
  relocate(decade, .after = year)
```

# Learning Something New: ANOVA & Tukey's Test

This week, I wanted to learn how to run an ANOVA and Tukey's test.   
The code was adapted from:  
[Tutorial for ANOVA and Tukey's Test](https://statdoe.com/one-way-anova-and-box-plot-in-r/)


## Preparing the data

For this data set, we will compare the average ratings of monster movies for the decades 1940, 1980, 1990, 2000, 2010 and 2020.
```{r}
anova_monster <- monster_movies %>% 
  # selecting only the decades of interest
  filter(decade >= 1980 | decade ==1940) %>% 
  # changing the decades into ordered factors
  mutate(decade = factor(decade,
                         levels=c("1940", "1980", "1990",
                                "2000", "2010", "2020"))) %>% 
  # sub-setting the data to only keep the columns of interests
  select(decade, average_rating)

head(anova_monster, 7)
```



```{r}
# running the anova test on the subset of data
anova <- aov(average_rating~decade, data = anova_monster)
summary.aov(anova)
```


```{r}
# Running the Tukey's test to compare the difference between groups
tukey <- TukeyHSD(anova)
print(tukey)
```


```{r}
# Compact Letter display: to show what groups are close to the other ones.
cld <- multcompLetters4(anova,tukey)
print(cld)
```


```{r}
# calculating the average and the quantiles of the average ratings by decades 
# this allows to sort the data but decreasing order of the average. 
# Quantiles will be used later

Tk<- anova_monster %>% 
  group_by(decade) %>% 
  summarize(avg = mean(average_rating, na.rm = T),
            quant = quantile(average_rating, probs = 0.75, na.rm = T)) %>% 
  arrange(desc(avg))
```


```{r}
# We pull the compact letter display from the cld data object and create a new column with the letters associated with the decades. 

# the quantiles will be used to plot the cld on the future box plots (y coordinates)

cld_df <- as.data.frame.list(cld[[1]])
Tk$cld <- cld_df$Letters
Tk
```



## Plotting 

```{r anova_ref, fig.width=12, fig.height=8, fig.path= "../outputs/"}
anova_monster %>% # plotting our subset data
  # we are interested in the decade and the average ratings
  ggplot(aes(decade, average_rating))+
  geom_boxplot(aes(fill = decade))+
  #plot the letters using the quantile values (y coordinates) and offsetting their positions so they don't overlap with the box plots.
  geom_text(data =Tk, aes(x= decade, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-1)+
  labs(title = "Distribution of the average movie ratings ",
       subtitle = "Comparison between selected decades",
       x = "Decade",
       y = "Average Rating",
       fill = "Decade") +
  theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        legend.position = "none",
        panel.background = element_rect(fill = "azure1"))+
  #custom the colors 
  scale_fill_ghibli_d("MarnieMedium1", direction =-1)
```

# Creating a function to generate the ANOVA and Tukey's test 
Obtaining the final plot is slightly annoying because of all the things that need to be calculated.   

The goal to create a function that can generate the ANOVA, Tukey's test and Compact Letter Display in fewer lines of code. 

```{r}
# creating the function anova_nunu because it's mine. 
# the arguments of the fonction will be our data set, the independant variable and the dependant variable of interest. 

anova_nunu <- function(data, indep_var, dep_var ) {

  # we want to make sure that the independent variable is a categorical variable. I set it up such that if the independent variable is not a factor, an error appears and stops the computation. 
  if (is.factor(data %>%  pull({{indep_var}})) ) {
    print("indep var is a factor")
  } else {
    stop("indep var is not a factor")
  }
  
  # same idea here for the dependent varible which has to be a numeric. Otherwise, an error shows. 
  if (is.numeric(data %>% pull({{dep_var}}))) {
    print("dep var is numeric")
  } else {
    stop("dep var is not numeric")
  }

  #Ok... So I couldn't figure out how to feed the data to aov() without having an error. I did resort to used chat GPT for that bit of code. 
  
  # I am not exactly sure what all the functions do, but the idea is the following. We take the column names from the data set, paste it so that it looks like "dep_var ~ indep var" (required input to the aov() function) and tell R to process that as a formula. 
  formula <- as.formula(paste(deparse(substitute(dep_var)), "~", deparse(substitute(indep_var))))
  
  # calculating the anova based on the formula above and with our data set.
  anova <- aov(formula, data = data)
  anova <<- anova # saves anova in the RStudio  environment
  
  # applying the Tukey's test
  tukey <- TukeyHSD(anova) 
  tukey <<- tukey # saves tukey in the RStudio  environment
  
  #generating the compact letter display (cld).
  cld <- multcompLetters4(anova,tukey)
  
  #generating a data set with the cld that can be plotted with the quantiles
  Tk<- data %>% 
    group_by({{indep_var}}) %>% 
    summarize(avg = mean({{dep_var}}, na.rm = T),
            quant = quantile({{dep_var}}, probs = 0.75, na.rm = T)) %>% 
    arrange(desc(avg))
  
  # pull the cld and adding them to the data set with the quantiles.
  cld_df <- as.data.frame.list(cld[[1]])
  Tk$cld <- cld_df$Letters
  
  # saves Tk in the RStudio  environment
  Tk <<- Tk  
  
  # tell the user that Tk has been saved and can be used for plotting
  print("Tk saved in environment" )
  return(Tk)

}

```

## Testing the Function

First, we will recreate the example plot from earlier with all the decades.

```{}
# This chunk shows an error because decade has not been converted to factor yet.  
This awesome because my function works ! 

anova_nunu(data = monster_movies, indep_var = decade, dep_var = average_rating)

```

```{}
Error in anova_nunu(data = monster_movies, indep_var = decade, dep_var = average_rating) : 
  indep var is not a factor
```


```{r}
# first we convert the decades to factors
monster_movies<- monster_movies %>% 
  mutate(decade = factor(decade)) %>% 
  drop_na()

# Apply the Nunu function
anova_nunu(data = monster_movies, indep_var = decade, dep_var = average_rating)
```

## Plotting 

As we can see, the function works and I was able to generate the cld from the ANOVA and Tukey's test for all the decades now. 
```{r anova_rating, fig.width=12, fig.height=8, fig.path= "../outputs/"}
monster_movies %>% 
  ggplot(aes(decade, average_rating))+
  geom_boxplot(aes(fill = decade))+
  geom_text(data =Tk, aes(x= decade, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(title = "Distribution of the average movie ratings ",
       subtitle = "Comparison between all decades",
       x = "Decade",
       y = "Average Rating",
       fill = "Decade") +
  theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        legend.position = "none",
        panel.background = element_rect(fill = "azure1")) 
# I couldn't add custom colors though, because the Ghibli palettes do not have enough colors to cover all the decades.
```

## More Examples
```{r anova_runtime, fig.width=12, fig.height=8, fig.path= "../outputs/"}
# Apply the Nunu function : Movie Run Time
anova_nunu(data = monster_movies, indep_var = decade, dep_var = runtime_minutes)

monster_movies %>% 
  ggplot(aes(decade, runtime_minutes))+
  geom_boxplot(aes(fill = decade))+
  geom_text(data =Tk, aes(x= decade, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(title = "Distribution of the movie run times",
       x = "Decade",
       y = "Run Time (min) [sqrt-transformed]",
       fill = "Decade") +
  coord_trans(y = "sqrt")+
  theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        legend.position = "none",
        panel.background = element_rect(fill = "azure1")) 

```

```{r anova_votes, fig.width=12, fig.height=8, fig.path= "../outputs/"}
# Apply the Nunu function: Number of votes 
anova_nunu(data = monster_movies, indep_var = decade, dep_var = num_votes)

monster_movies %>% 
  ggplot(aes(decade, num_votes))+
  geom_boxplot(aes(fill = decade))+
  geom_text(data =Tk, aes(x= decade, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(title = "Distribution of the number of votes",
       x = "Decade",
       y = "Number of votes (log10)",
       fill = "Decade") +
  coord_trans(y = "log10")+
  theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        legend.position = "none",
        panel.background = element_rect(fill = "azure1")) 

```

```{r anova_years, fig.width=12, fig.height=8, fig.path= "../outputs/"}
# Apply the Nunu function: Years 
#This plot is mostly for fun as it has no statistical meaning/relevance
anova_nunu(data = monster_movies, indep_var = decade, dep_var = year)

monster_movies %>% 
  ggplot(aes(decade, year))+
  geom_boxplot(aes(fill = decade))+
  geom_text(data =Tk, aes(x= decade, y = quant, label = cld),
            size = 4, vjust = -1, hjust =-0.5)+
   labs(title = "Distribution of the movie years",
       x = "Decade",
       y = "Year",
       fill = "Decade") +
  theme_bw() + 
  theme(plot.title = element_text(size=14, face = "bold"), 
        plot.subtitle = element_text(size=12),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        strip.text.x = element_text(size = 12, face = "bold"),
        legend.title=element_text(size=12, face = "bold"),
        legend.text=element_text(size=12),
        axis.text.x = element_text(angle= 45, hjust= 1),
        legend.position = "none",
        panel.background = element_rect(fill = "azure1")) 

```






