---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: Week 15 - Iterative Coding HW
output: 
  github_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(gt)
```

## Iterative Coding in Base R 
### Setting up the elements of the loop

Set the path to the files
```{r}
TPpath<- here("Week_15", "data", "homework")
```

Create a vector containing the files
```{r}
files <- dir(path = TPpath, pattern = ".csv")
files
```
Create an empty tibble with the varibles that we want to extract:

- filename: which CSV is the data coming from
- temp_mean: average temperature of the file
- temp_sd: standard deviation of temperature of the file
- light_mean: average light intensity of the file
- light_sd: standard deviation of light intensity of the file
```{r}
TPdata <- tibble(filename = rep(NA, length(files)),
                 temp_mean = rep(NA, length(files)),
                 temp_sd = rep(NA, length(files)),
                 light_mean = rep(NA, length(files)),
                 light_sd = rep(NA, length(files))
                 )
```

Testing if we can read a file using the paths and the directory set up above
```{r}
datatest<- read_csv(paste0(TPpath, "/", files[1]))
head(datatest, 10)
```

Making sure that we can calculate the mean based on a file 
```{r}
mean(datatest$Temp.C, na.rm = T)
```
### Setup the for loop

Making sure that loop works to read files
```{r}
for (i in 1:length(files)){
  rawdata <- read_csv(paste0(TPpath, "/", files[i] ))
  glimpse(rawdata)
}
```

Setup up the full for loop
```{r}
for (i in 1:length(files)){ # i in 1 to 4
  rawdata <- read_csv(paste0(TPpath, "/", files[i] )) # where to read the data
  
  TPdata$filename[i] <- files[i] # filename i in TP data is the file i
  
  TPdata$temp_mean[i] <- mean(rawdata$Temp.C, na.rm = TRUE) # calculate the quantities desired
  TPdata$temp_sd[i] <- sd(rawdata$Temp.C, na.rm = TRUE)
  TPdata$light_mean[i] <- mean(rawdata$Intensity.lux, na.rm = TRUE)
  TPdata$light_sd[i] <- sd(rawdata$Intensity.lux, na.rm = TRUE)
}

TPdata #read the populated tibble
```
### Returning a  Table Summary
```{r}
TPdata %>% gt(rowname_col = "filename") %>% 
  cols_label(temp_mean = "Avg Temperature (°C)",
             temp_sd = "Std Dev Temperature",
             light_mean = "Avg Light (Lux)",
             light_sd = "Std Dev Light")
```


## Iterative Coding with Tidyverse

### Setup to initiate the "looping"
Set the path to files and directory to access the files
```{r}
TPpath <- here("Week_15", "data", "homework")
files <- dir(path = TPpath, pattern = ".csv", full.names =  TRUE)
```
### Calculating the values with tidyerse
```{r}
files %>% 
  set_names() %>% 
  map_df(read_csv, .id = "filename") %>% # create a df with the first column
  group_by(filename) %>% # group the next values based on filename
  
  #calculating the desired quantities with tidyverse
  summarize(temp_mean = mean(Temp.C, na.rm = T),
            temp_sd = sd(Temp.C, na.rm = T),
            light_mean = mean(Intensity.lux, na.rm = T),
            light_sd = sd(Intensity.lux, na.rm = T)
            ) %>% 
  
  # change the filename to only keep the file name.csv and not the whole path directory
  mutate(filename = str_extract(filename, "[^/]+$")) %>% 
  
  #return a clean table summary with the values wanted
  gt(rowname_col = "filename") %>% 
  cols_label(temp_mean = "Avg Temperature (°C)",
             temp_sd = "Std Dev Temperature",
             light_mean = "Avg Light (Lux)",
             light_sd = "Std Dev Light")
```






