---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: Week 15 - Iterative coding
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

# Load libraries
```{r}
library(tidyverse)
library(here)
```

# For loops

Intial idea
```{r}
print(paste("The year is", 2000))
```

### Create a for loop
```{r}
years<- c(2015:2021)
```

```{r}
for (i in years) {
  print(paste("The year is", i))

}
```


### adding the outputs to a dataframe
```{r}
#Pre-allocate space for the df received in the output

year_data <- tibble(year= rep(NA, length(years)),
                    year_name= rep(NA, length(years)))
```


```{r}
for (i in 1:length(years)){
  year_data$year_name[i] <- paste("The year is", years[i])
  year_data$year[i] <- years[i]
}
```

### trouble shooting 

```{r}
year_data <- tibble(year= rep(NA, length(years)),
                    year_name= rep(NA, length(years)))

#We can use i=1 to test the for loop before running the whole for loop
i = 1

year_data$year_name[i] <- paste("The year is", years[i])
year_data$year[i] <- years[i]

year_data
```

# Load Data
```{r}
sal1 <- read_csv(here("Week_15", "data", "cond_data", "011521_CT316_1pcal.csv"))
#sal2 <- read_csv(here("Week_15", "data", "011621_CT316_1pcal.csv"))
#sal3 <- read_csv(here("Week_15", "data", "011721_CT354_1pcal.csv"))
```


```{r}
sal1
```

## Iterating on the csv files 

### List file directory
```{r}
# point to the location on the computer of the folder
CondPath<-here("Week_15", "data", "cond_data")
# list all the files in that path with a specific pattern
# In this case we are looking for everything that has a .csv in the filename
# you can use regex to be more specific if you are looking for certain patterns in filenames
files <- dir(path = CondPath, pattern = ".csv")
files
```
### Pre-allocate space
```{r}
# make an empty data frame that has one row for each file and 3 columns
 cond_data<-tibble(filename =  rep(NA, length(files)),  # column  for year
                   mean_temp = rep(NA, length(files)), # column for mean temp
                   mean_sal = rep(NA, length(files)), # column for mean salinity
                   ) # column name for the year name
cond_data
```


### For Loop
```#{r}
raw_data<-read_csv(paste0(CondPath,"/",files[1])) # test by reading in the first file and see if it works
head(raw_data)
```


Testing that we can calculate the mean in base R
```{r}
mean_temp<-mean(raw_data$Temperature, na.rm = TRUE) # calculate a mean
mean_temp
```

### Turning it into a for loop 

#### Can you read the file 
```{r}
for (i in length(files)){
  raw_data <- read_csv(paste0(CondPath,"/",files[i]))
  glimpse(raw_data)
}
```

```{r}
for (i in length(files)){
  raw_data <- read_csv(paste0(CondPath,"/",files[i]))
  #glimpse(raw_data)
  
  cond_data$filename[i] <- files[i]
}
cond_data
````

```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files 
raw_data<-read_csv(paste0(CondPath,"/",files[i]))
#glimpse(raw_data)
cond_data$filename[i]<-files[i]
cond_data$mean_temp[i]<-mean(raw_data$Temperature, na.rm =TRUE)
cond_data$mean_sal[i]<-mean(raw_data$Salinity, na.rm =TRUE)
} 
cond_data

```

# Map functions
## Use a canned function that already exist

Create a vector 1:10
```{r}
1:10
```

for each time 1:10 make a vector of 15 random numbers based on a normal distribution
```{r}
1:10 %>% 
  map(rnorm, n = 15)
```

Apply a canned function to the data 
```{r}
1:10 %>% 
  map(rnorm, n = 15) %>% 
  map_dbl(mean)
```

Can also use your own function
```{r}
1:10 %>% # list 1:10
  map(function(x) rnorm(15, x)) %>% # make your own function
  map_dbl(mean)
```
Use a formula when you want to change the arguments within the function
```{r}
1:10 %>%
  map(~ rnorm(15, .x)) %>% # changes the arguments inside the function, .x is kinda like the indexing elemen ~ [i]
  map_dbl(mean)
```




## Using Purrr to recreate files df that we made earlier 

```{r}
# point to the location on the computer of the folder
CondPath<-here("Week_15", "data", "cond_data")
files <- dir(path = CondPath,pattern = ".csv", full.names = TRUE)
files
```

OR 
```#{r}
files <- dir(path = CondPath,pattern = ".csv", full.names = TRUE)
#save the entire path name
files
```

## Read in the files 
```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") # map everything to a dataframe and put the id in a column called filename
data
```








