---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: 'Week 15 - Intro to Modeling '
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      messages = FALSE)
```

```{r}
library(here)
library(tidyverse)
library(palmerpenguins)
library(performance)
library(modelsummary)
library(tidymodels)
library(broom)
#library(flextable)
#library(see)

```

## Anatomy of a model (simple)
```{}
mod <- lm(y~x, data = df)

mod: name of the model
y~x: y as a function x 

```

Multiple regressions
```{}
mod <- lm(y~x1 + x2, data = df)
```

Interaction term 
```{}
mod <- lm(y~x1*x2, data = df)

* is computed as x1+x2+x1:x2 (x1 interacting with x2)
x1:x2 : interaction of x1 and x2
```


## Example

```{r}
peng_mod <- lm(bill_length_mm ~ bill_depth_mm*species, data = penguins)
```

### Checking the model assumption with performance

In lm, we assume that the data are meets 
- assumption of normality, i.e. residuals of the models are normal
- homogeneity of variance, 
- co-linearity....

```{r}
check_model(peng_mod)
```

### ANOVA table
- significant rel betwn  bill length and bill depth
- significant difference btwn species
- rel btwn bill depth and bill length change by spp

```{r}
anova(peng_mod)
```


### Coefficients (effect size) with error
gives:
- coefficients 
- intercepts 
- std error
- p-values
```{r}
summary(peng_mod)
```


## Broom: view results cleanly
Puts the results of the ANOVA and the coeff in a clean tibble
for export into Excel/CSV or plotting 
```{r}
coeffs <- tidy(peng_mod) #tidy() gives the ANOVA results
coeffs

```

### Glance() : R-squared, AICs... etc.
- r-squared
- adj. r-squared
- p-values...etc
```{r}
results <- glance(peng_mod)
results
```

### Residuals
Used for fitted values/fitted predictions, useful for plotting, std error and more
```{r}
resid_fitted <- augment(peng_mod)
resid_fitted
```


## Results un {modelsummary}

2 families of functions:
- modelsummary 
    - modelsummary: Regression tables with side-by-side models.
    - modelsummary_wide: Regression tables for categorical response models or grouped coefficients.
    - modelplot: Coefficient plots.
- datasummary
    - datasummary: Powerful tool to create (multi-level) cross-tabs and data summaries.
    - datasummary_balance: Balance tables with subgroup statistics and difference in means (aka “Table 1”).
    - datasummary_correlation: Correlation tables.
    - datasummary_skim: Quick overview (“skim”) of a dataset.
    - datasummary_df: Turn dataframes into nice tables with titles, notes, etc.

### Example
Produces a Word doc with both models' results
```{r}
# New model
peng_mod_noX<-lm(bill_length_mm ~ bill_depth_mm, data = penguins)
#Make a list of models and name them
models<-list("Model with interaction" = peng_mod,
             "Model with no interaction" = peng_mod_noX)
#Save the results as a .docx
modelsummary(models, output = here("Week_15","outputs","table.docx"))
```

## Model plot

```{r}
library(wesanderson)
```

```{r}
modelplot(models)+
  labs(x = "Coeffs", 
       y= "Term names")+
  scale_color_manual(values = wes_palette('Darjeeling1'))

```


# Running models on multiple parameters

Combine the purr, dplyr and broom packages. 
Create a set of lists and each have a dataset. 

for example purposes
```{r}
models <- penguins %>% 
  ungroup() %>% 
  nest(.by = species)

models

#turns the data in a bunch of smaller dfs by species
```

### Running a model for each of the species
```{r}
models <- penguins %>% 
  ungroup() %>% 
  nest(.by = species) %>% 
  mutate(fit = map(data, ~lm(bill_length_mm~body_mass_g, data = .)))

# here, set up a loop and run the model in the map() loop and mutate the results in a new column called "fit".

models

# this just computes the model but does not give the results yet
```
See the results of the model
```{r}
models$fit
```

### Calculating the results of the model
```{r}
results <- models %>% 
  mutate(coeffs = map (fit, tidy), #tidy = summary table 
         modelresults = map(fit, glance)) 

# creates a coeff column and apply the function tidy() to the fit column
# creates a modelresults column and apply glance() to fit column
results

```

### To see the results, we need to un-nest the list
```{r}
results <- models %>% 
  mutate(coeffs = map (fit, tidy), #tidy = summary table 
         modelresults = map(fit, glance)) %>% 
  select(species, coeffs,  modelresults) %>% 
  unnest(cols=c(coeffs,  modelresults), names_sep = "_") 

```

## {Tidymodels}
package that integrate tidyverse for modeling

1) specify the type of model 
```{r}
linear-reg()
```


