---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: Week 9 - Functional Programming
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(palmerpenguins)
library(PNWColors)
```


## Load the data
```{r}
income_mean <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_mean.csv')
```


## Definition
Factor: special character, used for categorical data. Usually alphabetical order but can be reordered. We need to change its "levels".  
Factors are stored as integers by R.

```{r}
fruits<-factor(c("Apple", "Grape", "Banana"))
fruits
```
### Booby-traps of Factors
**Normally:** if there is a typo in a column of numbers, R will read everything in as characters. When converting them back to a number, the rows with real characters will covert to NAs.  

**With factor:** It will display numbers associated with the levels of the factors and not show an NA when it's supposed to be.  

```{r}
test<-c("A", "1", "2")
as.numeric(test)
```

```{r}
test<-factor(test) # covert to factor
as.numeric(test)
# A converts to a 3. Not good!
```

## Forcats: family of function. _fct_XXX()_
We'll use starwars dataset.
```{r}
starwars
```

```{r}
starwars %>% 
  filter(!is.na(species)) %>% 
  count(species, sort = TRUE)

# 38 different species so we want to narrow it down to the most commonly occuring ones
```

```{r}
star_count <- starwars %>% 
  filter(!is.na(species)) %>% 
  mutate(species = fct_lump(species, n=3)) %>% 
   #converts spp to a factor and lumps spp with les than 3 indiv in an "other" category. 
  count(species) # without sort, it's alphabetical order.
star_count 
```
## Factors in plotting

Plotting as is:
```{r}
star_count %>% 
  ggplot(aes(x= species, y= n))+
  geom_col()

#what if we want to sort from high to low? 
```

Plotting in order of the n:
```{r}
star_count %>% 
  ggplot(aes(x= fct_reorder(species, n), y= n)) +
  geom_col()

#fct_reorder(X, Y), X: category to reorder, Y: reorder by variable Y
```
Making it descending order. Yay!
```{r}
star_count %>% 
  ggplot(aes(x= fct_reorder(species, n, .desc=T), y= n)) +
  geom_col()
```


## Better plotting with factors
Using the income_mean data
```{r}
glimpse(income_mean)
```
Tidying the data first
```{r}
total_income<-income_mean %>%
  group_by(year, income_quintile)%>%
  summarise(income_dollars_sum = sum(income_dollars))%>%
  mutate(income_quintile = factor(income_quintile)) # make it a factor
total_income
```

### Basic Plot
```{r}
total_income%>%
ggplot(aes(x = year, y = income_dollars_sum, color = income_quintile))+
  geom_line()

# the order of the income quintile is not in good order to read, we can rearrange it better
```

```{r}
total_income%>%
  ggplot(aes(x = year, y = income_dollars_sum, 
             color = fct_reorder2(income_quintile,year, income_dollars_sum)))+
   # fct_reorder2() reorders data based on 2 other variables
  geom_line()+
  labs(color = "income quantile")

# reorder the quintile by year first, then income sum


```

## Reordering factors in a dataset

```{r}
x1 <- factor(c("Jan", "Mar", "Apr", "Dec"))
x1

# Levels: Apr Dec Jan Mar | alphabetical order
```
But can be reordered:
```{r}
x1 <- factor(c("Jan", "Mar", "Apr", "Dec"), levels = c("Jan", "Mar", "Apr", "Dec"))
x1

#Levels: Jan Mar Apr Dec
```


## Subsetting data with factors

Grouping our species that have < 3 counts into "Other" we just filter them out.
```{r}
starwars_clean<-starwars %>% 
  filter(!is.na(species)) %>% # remove the NAs
  count(species, sort = TRUE) %>%
  mutate(species = factor(species)) %>% # make species a factor
  filter(n>3) # only keep species that have more than 3
starwars_clean
```
**BUT!** Levels of the factors are kept and display all categories
```{r}
levels(starwars_clean$species)
```

We need droplevels()

```{r}
starwars_clean<-starwars %>% 
  filter(!is.na(species)) %>% # remove the NAs
  count(species, sort = TRUE) %>%
  mutate(species = factor(species)) %>% # make species a factor
  filter(n>3) %>%  # only keep species that have more than 3
  droplevels() # can also use fct_drop()
starwars_clean
levels(starwars_clean$species)
```

## Recoding the levels
Relabiling the levels
```{r}
starwars_clean<-starwars %>% 
  filter(!is.na(species)) %>% # remove the NAs
  count(species, sort = TRUE) %>%
  mutate(species = factor(species)) %>% # make species a factor 
  filter(n>3)  %>% # only keep species that have more than 3 
  droplevels() %>% # drop extra levels 
  mutate(species = fct_recode(species, "Humanoid" = "Human"))
                               # Human is changed to Humanoid 
starwars_clean
```


## Awesome package of the day : gm to generate music

```{r}
library(gm)
```

```{r}

pitchesV<- list(c("E3","E4"), "B3", c("G3","E4"), "B3", 
                c("E3","E4"), "B3", c("G3","E4"), "B3",
                c("D3","D4"), "B3", c("G3","D4"), "B3",
                c("D3","D4"), "B3", c("G3","D4"), "B3",
                c("D3","D4"), "B3", c("F#3","D4"), "B3",
                c("D3","D4"), "B3", c("F#3","D4"), "B3",
                c("D3","D4"), "A3", c("F#3","D4"), "A3",
                c("D3","D4"), "A3", c("F#3","D4"), "A3")

line<- Line(pitchesV,
  durations = c("eighth", "eighth", "eighth", "eighth",
                "eighth", "eighth", "eighth", "eighth",
                "eighth", "eighth", "eighth", "eighth",
                "eighth", "eighth", "eighth", "eighth",
                "eighth", "eighth", "eighth", "eighth",
                "eighth", "eighth", "eighth", "eighth",
                "eighth", "eighth", "eighth", "eighth",
                "eighth", "eighth", "eighth", "eighth")) 
  

music <- 
  Music() +
  Meter(4, 4) +
  line +
  Tempo(100) +
  Clef("F") +
  Key(1, to = 1) +
  Pedal(1,4)+
  Pedal(5,8)+
  Pedal(9,12)+
  Pedal(13,16)+
  Pedal(17,20)+
  Pedal(21,24)+
  Pedal(25,28)+
  Pedal(29,32)
  

show(music)
```















