---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: Week 9 - Functional Programming
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Load Libraries
```{r}
library(tidyverse)
library(here)
library(palmerpenguins)
library(PNWColors)
```

## Why are function useful?

### Example: rescaling a dataframe
```{r}
df <- tibble(
  a= rnorm(10), #10 random values that are normally distributed
  b= rnorm(10),
  c= rnorm(10),
  d= rnorm(10),
)
head(df)
```


```{r}
df %>%
  mutate(a = (a-min(a, na.rm = TRUE))/(max(a, na.rm = TRUE)-min(a, na.rm = TRUE)))
# (a-min(a)) / (max(a)-min(a))
```


```{r}
df %>%
  mutate(a = (a-min(a, na.rm = TRUE))/(max(a, na.rm = TRUE)-min(a, na.rm = TRUE)),
         b = (b-min(b, na.rm = TRUE))/(max(b, na.rm = TRUE)-min(b, na.rm = TRUE)),
         c = (c-min(c, na.rm = TRUE))/(max(c, na.rm = TRUE)-min(c, na.rm = TRUE)),
         d = (d-min(d, na.rm = TRUE))/(max(d, na.rm = TRUE)-min(d, na.rm = TRUE)))

```
It's easy to make a mistake in copying the code.  
Fuctions can help to automate the process.

### Writing it in a function
```{r}
rescale01 <- function(x) {
  value <- (x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))
  return(value)
}
```

```{r}
df %>%
  mutate(a = rescale01(a),
         b = rescale01(b),
         c = rescale01(c),
         d = rescale01(d))
```

### Example: Farenheit to Celsius

Transforming the following in a function
```{}
temp_C <- (temp_F - 32)*5/9
```
Step 1: Name the function

```{r}
f_to_c <- function() {
  
}

```
Step 2: Put in the equation
```{r}
f_to_c <- function() {
  temp_C <- (temp_F - 32)*5/9
}

```
Step 3: decide of the arguments
```{r}
f_to_c <- function(temp_F) {
  temp_C <- (temp_F - 32)*5/9
  return(temp_C)
}

```
Step 4: Test it out

```{r}
f_to_c(350)
```

### Practice: C to K
```{r}
c_to_k <- function(temp_C) {
  temp_K <- (temp_C +273.15)
  return(temp_K)
}
```

```{r}
c_to_k(180)
```

## Making plots into a function

Color scales in a graph
```{r}
pal <- pnw_palette("Lake", 3, type = "discrete")
#choose the palette "Lake", only 3 of the colors, discrete colors

```

Basic graph
```{r}
ggplot(penguins, aes(x = body_mass_g, y = bill_length_mm, color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and another example of how to manually change the legend title for colors
  theme_bw()
```

### Making plots into functions

The skeleton of the function: what arguments should be included
``` {}
myplot<-function(){
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(penguins, aes(x = body_mass_g, y = bill_length_mm, color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ 
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
```

Swap the defined values to arguments in function
```{r}
myplot<-function(data, x, y){
pal<-pnw_palette("Lake",3, type = "discrete") 
#specify the df, x and y values for future plotting
ggplot(data, aes(x = x, y =y , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ 
  scale_color_manual("Island", values=pal)+   
  theme_bw()
}
```

Because x and y are arguments from within the dataset. We need the specify that to R.  
We do that by using x = {{x}}, y ={{y}}.
```{r}
myplot<-function(data, x, y){
pal<-pnw_palette("Lake",3, type = "discrete") 
#specify the df, x and y values for future plotting
ggplot(data, aes(x = {{x}}, y ={{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ 
  scale_color_manual("Island", values=pal)+   
  theme_bw()
}
```
Note: No need to specify "return" because ggplot object not assign to a
variable

```{r}
myplot(data = penguins, x = body_mass_g, y = bill_length_mm)
#easy plotting with no need to type the same code all the time
```

### Adding default parameters in the function
The default df is the penguins
```{r}
myplot<-function(data = penguins, x, y){
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
```

```{r}
myplot(x = body_mass_g, y = bill_length_mm)
```
### Can layer the plot as normal because output is a ggplot object 

```{r}
myplot(x = body_mass_g, y = flipper_length_mm)+
  labs(x = "Body mass (g)",
       y = "Flipper length (mm)")

#and can add more layers 
```

## Flexibility in the function: ifelse 
```{r}
a <- 4
b <- 5
```

```{r}
if (a>b) {
  f <-20
} else {
      f <- 10
}

f # f is equal to 10. 

# example of conditional statements
# can be used in functions
```

## Conditional functions
Whether or not to add a linear regression

### Backbone:
```{}
myplot<-function(data = penguins, x, y ,lines=TRUE ){ # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete") 
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ 
  scale_color_manual("Island", values=pal)+ 
  theme_bw()
}
```


```{r}
myplot<-function(data = penguins, x, y, lines=TRUE ){ # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete")  

#Add the conditions: 
if(lines==TRUE){
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model IF lines = TRUE
  scale_color_manual("Island", values=pal)+   
  theme_bw()
}

# Add alternative if contidition is false:
else{
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  scale_color_manual("Island", values=pal)+ 
  theme_bw()
}
}
```

Test

```{r}
#Lines = TRUE
myplot(x = body_mass_g, y = flipper_length_mm)
```

```{r}
#Lines = FALSE
myplot(x = body_mass_g, y = flipper_length_mm, lines = FALSE)
```












