---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: Week 12 - R Shiny
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

# Load Libraries
```{r}
library(tidyverse)
library(reprex) # reproducible example
library(datapasta) # copy and paste
library(styler) # needs to be load WITH datapasta
library(here)
```

# Reprex:

Allows to format R session it in a way that's easily shared on help website

Copy the snippet of code that causes issues -> Addins -> Render reprex -> "on the clipboard" -> "Github or Stackoverflow" -> "Append Session info"
```{r}
mpg %>% 
  ggplot(aes(x = displ, y = hwy))%>%
  geom_point(aes(color = class))
```


# Datapasta
Can copy and paste a piece of data frame and transform it in a tibble

Copy the head of your **CSV file** and paste it as a tribble with data paster
```{r}

data <- tibble::tribble(
    ~lat,    ~long, ~star_no,
  33.548, -117.805,      10L,
  35.534, -121.083,       1L,
  39.503, -123.743,      25L,
  32.863,  -117.24,      22L,
   33.46, -117.671,       8L,
  33.548, -117.805,       3L
  )

```


```{r}
iris %>% 
  ggplot(aes(x= Speal.Length, y = Petal.Width))
  geom_point(aes(color= species))
```

```{r}
ggplot() +
  geom_point(data = penguins, aes(x = bill_length_mm, y = species)) +
  labs(title = "PENGUINS",
       subtitle = "HOW LONG ARE THEIR BILLS")
```


