---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: Week 8 - Advanced Plotting
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```


## Load the Libraries
```{r}
library(tidyverse)
library(here)
library(patchwork)
library(ggrepel)
library(gganimate)
library(magick)
library(palmerpenguins)

```

## Patchwork
Brings plots together: combine and arrange plots.

### Example: 
```{r}
# plot 1
p1<-penguins %>%
  ggplot(aes(x = body_mass_g, 
             y = bill_length_mm, 
             color = species))+
  geom_point()
p1
```

```{r}
# plot 2
p2<-penguins %>%
  ggplot(aes(x = sex, 
             y = body_mass_g, 
             color = species))+
  geom_jitter(width = 0.2)
p2
```
To use patchwork, you just math signs to combine plots. Extremely easy. 
```{r}
#Simple plot combination 
p1+p2 # Plots side by side
p1/p2 # plots on top of each other
```
 #### Combining legends
```{r}
# combining legends together since they are the same
p1+p2 +
  plot_layout(guides = "collect") # guides = legends, basically
```

#### Add annotation to the graphs
```{r}
p1+p2 +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "A") # labels plots A-Z
```

Another way 
```{r}
p1/p2 +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "1") # labels plots A-Z
```


## ggrepel
Using mtcars package to show the functionalities of ggrepel 

```{r}
view(mtcars)
```

Example: Simple Plot
```{r}
ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   #extracts the labels from rows
                   label = rownames(mtcars))) + 
  geom_text() + # creates a text label
  geom_point(color = 'red', size = 2)
```

Repeling the labels so it's more visible 
```{r}
ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   label = rownames(mtcars))) + 
  geom_text_repel() + # creates a text label that don't overlap
  geom_point(color = 'red', size = 2)
```

Other way to make labels
```{r}
ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   label = rownames(mtcars))) + 
  geom_label_repel() + # creates a label stickers that don't overlap
  geom_point(color = 'red', size = 2)
```

## gganimate
Make animated figures 

Example: Static Plot
```{r}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point()
```

 ### Adding transition for movement
```{r}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(
     year, # what are we animating by
     transition_length = 2, #The relative length of the transition.
     state_length = 3 # The length of the pause between transitions
   )
```

### Adding aesthetics to transitions 
```{r}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(
     year, 
     transition_length = 2, 
     state_length = 1 
   ) +
  ease_aes("circular-in-out") #specifies the style of transition
```

### Add labels to the transitioned plots
```{r}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(
     year, 
     transition_length = 2, 
     state_length = 1 
   ) +
  ease_aes("circular-in-out") + #specifies the style of transition
  labs(title = "Year: {closest_state}") 
```

##### changing the plot and messing around
```{r}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(
     sex, 
     transition_length = 2, 
     state_length = 1 
   ) +
  ease_aes("circular-in-out") + 
  labs(title = "Sex: {closest_state}")
```

### Saving the animated plot as a gif
```{r}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(
     year, # what are we animating by
     transition_length = 2, #The relative length of the transition.
     state_length = 1 # The length of the pause between transitions
   )+
  ease_aes("sine-in-out") +
  labs(title = 'Year: {closest_state}') +
  anim_save(here("Week_08","outputs","mypengiungif.gif"))

```


## Magick
Advanced image processing

Adding a penguin to the plot
```{r}
penguin<-image_read("https://pngimg.com/uploads/penguin/pinguin_PNG9.png")
penguin
```

1. Need to create a plot and save in the outputs folder
```{r}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point()

ggsave(here("Week_08","outputs","penguinplot.png"))
```

2. Read the plot png back in R, then add the image
```{r}
#read the files in layers: 1st file is background, last file is foreground
penplot<-image_read(here("Week_08","outputs","penguinplot.png"))
out <- image_composite(penplot, penguin, offset = "+20+70") # offset: where the penguin should be 
out
```


3. can also add gifs on the plot 
```{r}
# Read in a penguin gif
pengif<-image_read("https://media3.giphy.com/media/H4uE6w9G1uK4M/giphy.gif")
outgif <- image_composite(penplot, pengif, gravity = "center") 
      # gravity = where does it pull 
animation <- image_animate(outgif, fps = 10, optimize = TRUE)
      # fps= specify the quality of the animated gif. fps 10 for low quality
      # less space
animation
```






