---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: Week 7 - Maps
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```


## Load the Libraries

```{r}
library(tidyverse)
library(here)
library(ggspatial)
library(ggmap)
```


## Add API Key to R

Google API: Need to unable Google API on static maps 
https://www.littlemissdata.com/blog/maps
```{r}
#ggmap::register_google(key = "Key")
```

Also add a Stadia Maps API 
```{r}
#register_stadiamaps("Key")
```

GGmaps Git hub : https://github.com/dkahle/ggmap

## Load the data
```{r}
chemdata <- read_csv(here("Badplot_Goodplot", "data", "chemicaldata_maunalua.csv"))

```

## Plotting with ggmap()

### Get base map from ggmap

```{r}
#get the map data
oahu <- get_map("Oahu")

#allows to plot data
ggmap(oahu)
```

### Focusing on specific coordinates for Wailupe

Centering the map around a coordinate
```{r}
WP <- data.frame(lon = -157.7621, lat = 21.27427)

map1 <- get_map(WP)

ggmap(map1)
```

### Zooming on a location

get_map(..., zoom = X)  
zoom is between 3 (zoomed out) and 20 (zoomed in)  
```{r}
map1 <- get_map(WP, zoom = 16)

ggmap(map1)
```
### Change the map type 
get_map(..., maptype = X)

```{r}
map1 <- get_map(WP, zoom = 16, maptype= "satellite")

ggmap(map1)
```

Fun map type : 
```{r}
map1 <- get_map(WP, zoom = 16, maptype= "stamen_watercolor", source = "stadia")

ggmap(map1)
```


### Plotting

Behaves like a ggplot 
```{r}
map1 <- get_map(WP, zoom = 17, maptype= "satellite")

ggmap(map1)+
  geom_point(data = chemdata,
             aes(x= Long, y= Lat, color = Salinity),
             size= 4)+
  scale_color_viridis_c()
```


```{r}
ggmap(map1)+
  geom_point(data = chemdata,
             aes(x= Long, y= Lat, color = Salinity),
             size= 4)+
  scale_color_viridis_c() +
  annotation_scale( bar_cols = c("red", "white"),
                    location = "bl")+ # puts scale bar on the bottom left (bl)
                                      # change colors to red and white
  annotation_north_arrow(location = "tl")+ # add a north arrow, top left (tl)
                              # many more editing options 
  coord_sf(crs = 4326) # specify the coordinate system to adjust the scale bar
                       # crs specifies the system: GPS (WGS84)

```


## Geocoding 
when you don't know the exact lat and long

```{r}
geocode("the White House")
geocode("University of Hawaii at Manoa")
```







