---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: Week 7 - Maps
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```


## Load the Libraries

```{r}
library(tidyverse)
library(here)

library(maps)
library(mapproj)
library(mapdata)

#sf, ggmap, tmap, leaflet are other packages that can be useful for GIS purposes
```

```{r}
popdata <- read_csv(here("Week_07", "data", "CApopdata.csv"))

stars <- read_csv(here("Week_07", "data", "stars.csv"))
```


## {maps} package
comes with some polygons and base maps. 
Last updated 2022.

```{r}
world <- map_data("world")
usa <- map_data("usa")
italy <- map_data("italy")

#the package is very US centric though
states <- map_data("state")
counties <- map_data("county")

```

## Structure of the data

long: longitude things West of the prime meridian is negative
lat : latitude  
order: order in which ggplot should connect the dots
region and subregion: tell what region a set of points surrounds
group : 


## Basic map
```{r}
ggplot() +
  geom_polygon(data = world, aes(x=long, y=lat, group = group,
                                 fill = region), # colot by regions
                                 color = "black") + #black outlines on the map
  guides(fill = FALSE) + # remove the legend because too many
  theme_minimal()+
  theme(panel.background = element_rect(fill = "lightblue"))
  
```

## Projections

```{r}
ggplot() +
  geom_polygon(data = world, aes(x=long, y=lat, group = group,
                                 fill = region), # colot by regions
                                 color = "black") + #black outlines on the map
  guides(fill = FALSE) + # remove the legend because too many
  theme_minimal()+
  theme(panel.background = element_rect(fill = "lightblue"))+

#change the projection of the map
  coord_map(projection = "mercator",
            xlim = (c(-180,180)))
  
```



```{r}
ggplot() +
  geom_polygon(data = world, aes(x=long, y=lat, group = group,
                                 fill = region), # colot by regions
                                 color = "black") + #black outlines on the map
  guides(fill = FALSE) + # remove the legend because too many
  theme_minimal()+
  theme(panel.background = element_rect(fill = "lightblue"))+
  coord_map(projection = "sinusoidal",
            xlim = (c(-180,180)))
  
```


## Map just of California

```{r}
head(states)
#map data are simple dataframes 
#don't overthink it
```

```{r}
# from the states data, just select the data relevant to CA
CA_data <- states %>% 
  filter(region == "california")
```

```{r}
ggplot (data = CA_data)+
  geom_polygon(aes(x= long, y = lat, group=group), 
               color = "lightblue") +
  coord_map(projection = "sinusoidal")
```


##Adding layers to the map

### Population density of CA
Joining data 
```{r}
CApop_county <- popdata %>%
  select("subregion" = County, Population)  %>% # rename the county col
  inner_join(counties) %>%
  filter(region == "california") # some counties have same names in other states

head(CApop_county)
## Joining with by = join_by(subregion)
# Check out the jannitor package ! Very useful


```

Map it with ggplot
```{r}
ggplot (data = CApop_county)+
  geom_polygon(aes(x= long, y = lat, group=group, 
                   fill = Population), #now we can see the pop density on the map
               color = "lightblue") +
  coord_map(projection = "sinusoidal")+
  scale_fill_gradient(trans = "log10")
```


### Add the seastar data on top of the CA map

```{r}
ggplot ()+
  geom_polygon(data = CApop_county,
               aes(x= long, y = lat, group=group, 
                   fill = Population), #now we can see the pop density on the map
               color = "lightblue") +
  geom_point(data = stars, 
           aes(x=long, y=lat,
               size = star_no))+
  coord_map()+
  scale_fill_gradient(trans = "log10") +
  labs(size = "# stars/m2") 

ggsave(here("Week_07","outputs","CApop.pdf"))

```



## Fun packages ggdogs
